Independent Component Analysis

### 2022-04-29 paper.tex
edited some part in mathematical formulation

As we discussed previously, we have the freedom to whiten the observed signals and we will denote the whitened observed data as $\tilde{\bf{x}}$, where each component is $\tilde{x_{j}} = \frac{x_{j}-\mathbb{E}[x_{j}]}{\sigma_{x,j}}$. The skewness, kurtosis and excess kutosis are defined as $$ \gamma(x) := \mathbb{E}[\tilde{x}^{3}], \beta(x) := \mathbb{E}[\tilde{x}^{4}], \kappa(x) := \beta(x)-3 $$.
The objective function is a convex combination of the squared third and fourth cumulants defined as follows,
\begin{equation}
L_{\alpha}(\bf{u}) := \alpha \gamma^{2}(\bf{u}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{u}^{t}\tilde{\bf{x}})
\end{equation},
where $\alpha \in [0, 1]$ and $\bf{u}^{t}\bf{u}$$ = 1$.

@@

We can first consider the basic idea of minimizing mutual information, which is defined as the KL divergence of the multivariate distribution of the observed data $\bf{x}$ and product of all its marginal distributions, i.e. 
\begin{equation}
I(\bf{x}) = \sum_{\bf{x}}p(\bf{x})\log\frac{p(\bf{x})}{\prod_{i}p_{i}(x_{i})}
\end{equation}. $I(\bf{x})$ is non-negative and is 0 if an only if the $\bf{x}$ are independent. Within this framework, several different approaches have been formulated in history. One of them is a key finding in neural networks, that is maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$. Maximizing the output entropy is to maximize $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$. In a neural network, there is a non-linear activation function that generates the output $\bf{y} = g(\bf{u})$. This nonlinearity indeed brings the information of higher order statistics into the optimization process.

@@

\section{How to check the robustness and }

1) check if the performance matrix is identity (Te-won Lee, p. 51)
\begin{align}
W A &= I.
\end{align}

2) check how fast w changes (ICA_projectionPursuit1612.05445.pdf, p. 12)

@@

\section{what if two of the sources are dependent}
s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
s4 = s1 + s2
s5 = s1 - s2

>>>>

As we discussed previously, we have the freedom to whiten the observed signals and we will denote the whitened observed data as $\tilde{\bf{x}}$, where each component is $\tilde{x_{j}} = \frac{x_{j}-\mathbb{E}[x_{j}]}{\sigma_{x,j}}$. The skewness, kurtosis and excess kutosis are defined as $$ \gamma(x) := \mathbb{E}[\tilde{x}^{3}], \beta(x) := \mathbb{E}[\tilde{x}^{4}], \kappa(x) := \beta(x)-3 $$.
The objective function is a convex combination of the squared third and fourth cumulants defined as follows,
\begin{equation}
L_{\alpha}(\bf{u}) := \alpha \gamma^{2}(\bf{w}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{w}^{t}\tilde{\bf{x}})
\end{equation},
where $\alpha \in [0, 1]$ and $\bf{w}^{t}\bf{w}$$ = 1$.

@@

We can first consider the basic idea of minimizing mutual information, which is defined as the KL divergence of the multivariate distribution of the observed data $\bf{x}$ and product of all its marginal distributions, i.e. 
\begin{equation}
I(\bf{x}) = \sum_{\bf{x}}p(\bf{x})\log\frac{p(\bf{x})}{\prod_{i}p_{i}(x_{i})}
\end{equation}. $I(\bf{x})$ is non-negative and is 0 if and only if the $\bf{x}$ are independent. Within this framework, several different information theoretic approaches have been formulated in history.

One of them is a key finding in neural networks, that is, maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$. Maximizing the output entropy is to maximize $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$. The maximum is obtained if marginal distribution of each $y_{i}$ is uniform and independent. This implies that the activation function in the neural network $y_{i} = g(\bf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$.

In a neural network, there is a non-linear activation function that generates the output $\bf{y} = g(\bf{u})$. This nonlinearity indeed brings the information of higher order statistics into the optimization process.

Another way to formulate this is to maximize the negentropy, which is the KL divergence between $p(\bf{W}\bf{x})$ and Gaussian distribution $p_{G}(\bf{W}\bf{x})$.

Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution.

@@

\section{How to check the robustness and }
% Amari metric, ref "ELements of Statistical Learning", p.570

%1) check if the performance matrix is identity (Te-won Lee, p. 51)
\begin{align}
W A &= I.
\end{align}

%2) check how fast w changes (ICA_projectionPursuit1612.05445.pdf, p. 12)

@@

\section{what if two of the sources are dependent}
%s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
%s4 = s1 + s2
%s5 = s1 - s2
